\documentclass[10pt, a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{ifpdf}

\title{Cognition, Report exercise 1}
\date{}
\author{Oscar Olsson, 900311 - 0690\\ Tommy Olofsson, 900814 - 4553\\
	Erik Westrup, 901021 - 1192\\ Erik Jansson, 901202 - 4114}
\begin{document}
\maketitle
\begin{center}
Group: 4 \\
Teacher: Thomas Strandberg.
\end{center}
\newpage

\section{Preamble}
The purpose of this laboratory exercise was to familiarize us with a variety of cognitive phenomena. We were subject to a number of demonstrations and active experiments.

\section{Inattention \& change blindness}
\subsection{Description}
To illustrate \emph{change blindness} we watched a video clip\footnote{http://www.youtube.com/watch?v=voAntzB7EwE} showing a man and a woman behind a desk, performing a card trick with a blue backed deck. The woman chose a random card from the deck and shows it to the camera, stating its color and number out loud. She then gave it back to the man, who put it back in the deck. The man then spreads the deck out, face up, on the table and picks out the previously chosen card. When the man then turned the deck over, all the other cards had a red back except for the one the woman had chosen. What also changed was the color of the background cape, both persons shirts and the table-cloth. These changes were coordinated with the camera movement so that the table-cloth could quickly be changed when the camera focused right above the table on the card etc.

While the audience focused on following the changes of card colors they missed all the other changes even though they are really apparent. Simons \& Levin have studies this phenomena in the late nineties with experiments on psychology students and people at the campus. \footnote{http://www.sierratrialandopinion.com/papers/ChangeBlindnessPaper.rtf}

Their first experiment illustrating the concept was set at a University campus where a man asks a random pedestrian for the direction to a building. \footnote{http://www.youtube.com/watch?v=FWSxSQsspiQ}  Two people caring a door then unlordly interrupts them by walking between them. When they walked through the person asking for direction changed position with one of the door carriers with out the helpful man noticing and he continues describing the way. Here again, by concentrating on a task, one fails to notice a change.

\subsection{Discussion}
In this exercise we observed the phenomena called \emph{change blindness}. In the color card trick exercise we all missed the changes in the surroundings due to focusing on discovering the secret behind the card trick itself. This was surprising to us, we did not expect anything but the card trick to occur. 

As discussed in “Unconscious Transference” As An Instance of Change Blindness \footnote{“Unconscious Transference” As An Instance of Change Blindness, http://www.sierratrialandopinion.com/papers/ChangeBlindnessPaper.rtf} eye witness testimonials might be more unreliable than previously considered due to these phenomena. From this we learn that describing something from memory other than what we focused on at the time can not be considered reliable. 

\section{Exercise 2, The MacGurk Effect}
\subsection{Description}
The McGurk effect is about the way our perception of speech differ depending on visual stimuli.

\subsection{The experiment}
The experiment was quite simple and straight-forward.  We got to view a video. First we were instructed to look down and just listen to the voice.  We heard nothing more than a man saying \emph{Ba, ba, ba, ba}.  Then we got to view the same video once again but this time we were allowed to look up at the screen.  The video was depicting the lower half of a face.  This time it sounded, for some of us,  like the man said something like \emph{ba, la, la, va} It turns out that we were actually hearing \emph{the same sound} but seeing different mouth-movements.

\subsection{Thoughts}
Although only half the group did actually notice the McGurk effect we can imagine the effect having some real-world significance. This might very well be the reason for people prefer video-conferencing and such when communication with friends and colleagues online i.e. Skype.

\section{Exercise 4, Memory demonstration}
\subsection{Description}
The exercise session began with the teacher handing out a small note to half of the class, containing the text \emph{"To wash clothes"} written in Swedish ("Att tvätta kläder"). The other half of the class was left in the dark without any note at all. He then began reading a text with a generic description of how to wash clothes and told us it was not that important and we should not pay that much attention to it. After this the exercise session continued with other topics and 30-45 minutes later the entire class were asked to recall as much of the text as possible could and write it down.

The result was, perhaps surprisingly, that the persons who were shown the heading \emph{"To wash clothes"} generally had easier to remember more pieces of the read text.

\subsection{Discussion}
The point of this exercise was to point out that the presence of a heading makes it easier to memorize a piece of text since the brain can associate the text with a short sentence. This effect is even greater if the person is familiar with the subject at hand\footnote{Headings as Memory Facilitators: The Importance of Prior Knowledge, http://soe.usd.edu/amandamcdonald/4.pdf}, and, since washing clothes is a familiar task for most people (hopefully), the group of people that had something to relate the text to, i.e. the heading, had much easier to memorize it.

However, we did not noticed this effect in our group. Since the teacher pointed out that this shouldn't be payed much attention to, none the members of our group even tried to memorize the text.

\subsection{Exercise 5, The Stroop Effect}
\subsubsection{Description}
The Stroop effect describes the peculiar fact that people have to devote a lot more conscious thought to say the color of a printed word compared to just reading the word. Particularly so if the word is the name of a color printed in another one, creating a mental clash.

The material for this experiment included just three cards.  Each card had about 15 words, one on each line, printed in a different color.  The words on the first card were all gibberish. This made it easy to determine the color with few errors and little stuttering. 
The second card was quite easy once we realized that the words were printed in the 'right' color. It cut the time by a couple of seconds and we had even fewer errors and less stuttering.

The third card was the hard one. The words were all names of different colors, but they were printed in a different one. For example, we might have had the word \emph{green} printed in red. This made it really hard to say the color of the print rather than reading the word itself.  We hesitated a lot and it took about twice the time to go through this card compared to the other cards. 

\subsubsection{Discussion}
Since we allready knew about this effect we weren't surprised to a see significant difficulty when the words and the color did not match. The Stroop test can be combined with other methods to find e.g. attention disorders such as ADHD (allthough with questionable results). \footnote{Stroop Interference and Attention-Deficit/Hyperactivity Disorder, Review and Meta-Analysis, http://psycnet.apa.org/journals/neu/21/2/251.pdf}. We have tried to find nautral occurences of the Stroop effect but failed to do so.\\[5mm]

\begin{tabular}{l l l l}
  Card & Characteristics & Mean time (s) & Mean number of errors \\
  \hline
  1 & non-words & 11.25 & 1.25 \\
  2 & colors right & 8.5 & 0.25 \\
  3 & colors wrong & 19.25 & 2.25 \\
\end{tabular}
\section{Exercise 6, Communication}

\subsection{Description}
This exercise was performed to illustrate how communication was affected by different limitations. We were given two copies of 18 figures divided on three separate pages. In one copy the figures was numbered from one through six on each page and in the other copy the figures were not numbered. One of us, the explainer, received the numbered figures and one of us, the responder, received the unnumbered figures. The receiver was asked to help the responder number the figures according to the receivers figures as fast as possible. The responder and receiver were not allowed to see each others figures nor draw figures on paper or in the air. The task was performed three times, each with a new set of six figures, with different limitations.
The limitations were:
\begin{enumerate}
\item The responder may not speak.
\item The responder and receiver may not maintain visual contact.
\item No limitations.
\end{enumerate}

\subsection{Discussion}
We performed a slightly modified version of the exercise due to a missed detail of the instructions (you were not allowed to draw figures). In all three tests we found no problem in communication. In the first part the explainer used speech and the responder confirmed with hand gestures. In the second part the responder confirmed with speech. I the last part explainer communicated with drawing of figures and the responder used speech to confirm. Surprisingly the best result was achieved when no visual contact was allowed. This might be due to a difference in difficulties in explaining the figures. From this experiment we can conclude that it is easier to communicate with speeach with no visual contact than only visual contact. This confirmes the comon notion that speech is more powerful than visualisations. For example it is more important to implement a good speech feature than a video feature in a conference system.

%\bibliographystyle{IEEEtran}
%\bibliography{refs}
\end{document}
